---
title: "Homework 2"
author: Jinghan Zhang
date: September 24, 2020
output: github_document
---

## Problem 1

```{r setup}
library(tidyverse)
library(readxl)
```

Read the Mr. Trashwheel dataset.

```{r data cleaning}
trashwheel_df <-
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    sport_balls = as.integer(sports_balls)
  )
```

Read precipitation data.

```{r precipitation tables}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip =1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018) %>%
  relocate(year)

precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip =1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017) %>%
  relocate(year)

month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  ) 

precip_df = bind_rows(precip_2018, precip_2017) 
left_join(precip_df, month_df, by = "month") 
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, including some specific types of trash. There are a total of `r nrow(trashwheel_df)` rows or observations in our final dataset. Additional data sheets include month precipitation data. The total precipitation in 2018 was `r sum(precip_2018$total)` inches and there was a median of `r median(pull(trashwheel_df, sports_balls), year == 2017, na.rm = TRUE)` sports balls in a dumpster in 2017.

## Problem 2

Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful). Reformat data so that route number and route name are distinct variables.

```{r setup for nyc transit}
nyc_transit <- read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  unite(route, route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11, sep = "/", na.rm = TRUE) %>%
  select('line', 'station_name', 'station_latitude', 'station_longitude', 'route', 'entry', 'vending', 'entrance_type', 'ada') %>%
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))

```

This dataset contains information about NYC's subway entrances and exits. Each row of the raw dataset refers to an individual station entrance or exit. There are a total of `r  nrow(nyc_transit)` rows or individual station entrances/exits. The raw dataset includes `r ncol(nyc_transit)` columns that detail the line, station name, station longitude/latitude, routes, entrance type, whether entry was allowed, presence of ticket vending, and whether it was ADA compliant. The data has been cleaned, with extraneous variables omitted and the entry variable converted from character to a logical variable. The data is not yet tidy, as there are repeated observations. There are `r select(nyc_transit, station_name, route) %>% distinct() %>% nrow()` distinct stations, with `r select(nyc_transit, station_name, route) %>% distinct() %>% dplyr::filter(grepl("A", route, fixed = TRUE)) %>% nrow()` stations specifically serving the A train, and `r select(nyc_transit, station_name, route, ada) %>% distinct() %>% dplyr::filter(grepl("A", route, fixed = TRUE)) %>% dplyr::filter(ada) %>% nrow()` of those stations are ADA compliant. However, there are `r select(nyc_transit, station_name, route, ada) %>% distinct() %>% dplyr::filter(ada) %>% nrow()` total ADA compliant stations. `r select(nyc_transit, station_name, route, vending, entry) %>% dplyr::filter(vending=="NO") %>% dplyr::filter(entry) %>% nrow() / select(nyc_transit, station_name, route, vending) %>% dplyr::filter(vending=="NO") %>% nrow() *100`% of the station entrances/exits without vending allow entrance.

## Problem 3

First, clean the data in pols-month.csv. Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.

Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.

Join the datasets by merging snp into pols, and merging unemployment into the result.

```{r setup for FiveThirtyEight data}
pols_month <- read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, c("year", "month", "day"), sep="-") %>%
  mutate(month = recode(month,
    "01" = "January",
    "02" = "February",
    "03" = "March",
    "04" = "April",
    "05" = "May",
    "06" = "June",
    "07" = "July",
    "08" = "August",
    "09" = "September",
    "10" = "October",
    "11" = "November",
    "12" = "December",
  )) %>%
  mutate(president = ifelse(prez_dem == 1, "dem", "gop")) %>%
  select(-day, -prez_gop, -prez_dem)

snp <- read_csv("./data/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, c("month", "day", "year"), sep="/") %>%
  mutate(month = recode(month,
    "1" = "January",
    "2" = "February",
    "3" = "March",
    "4" = "April",
    "5" = "May",
    "6" = "June",
    "7" = "July",
    "8" = "August",
    "9" = "September",
    "10" = "October",
    "11" = "November",
    "12" = "December",
  )) %>%
  select(year, month, -day, close)

employment <- read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(!year, names_to = "month", values_to = "unemployment") %>%
  mutate(month = recode(month,
    "jan" = "January",
    "feb" = "February",
    "mar" = "March",
    "apr" = "April",
    "may" = "May",
    "jun" = "June",
    "jul" = "July",
    "aug" = "August",
    "sep" = "September",
    "oct" = "October",
    "nov" = "November",
    "dec" = "December",
  ))

merged = merge(pols_month, snp) %>%
  merge(employment)

```

The first dataset pols_month has `r ncol(pols_month)` columns/variables, and `r nrow(pols_month)` rows/observations. It gives the breakdown of political representatives in the senate, house, governor, and presidential positions by republican versus democratic party per month starting from `r min(pull(pols_month, year))` to `r max(pull(pols_month, year))`. The actual variables are named `r colnames(pols_month)`.

The second dataset snp has `r ncol(snp)` columns/variables, and `r nrow(snp)` rows/observations. It gives what appears to be monthly market closing data starting from `r max(pull(snp, year))` and then going back until `r min(pull(snp, year))`. The actual variables are named `r colnames(snp)`.

The third dataset employment has `r ncol(employment)` columns/variables, and `r nrow(employment)` rows/observations. It gives yearly unemployment data by month from `r min(pull(employment, year))` to `r max(pull(employment, year))`. The actual variables are named `r colnames(employment)`.

The final dataset merged consists of `r nrow(merged)` rows where each row is data for a month, starting from `r min(pull(merged, year))` and ending at `r max(pull(merged, year))`. The data has `r ncol(merged)` columns consisting of the variables: `r colnames(merged)`.
